---
title: "nasapower: NASA-POWER Agroclimatology Data from R"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nasapower)
```

[![Travis-CI Build Status](https://travis-ci.org/adamhsparks/nasapower.svg?branch=master)](https://travis-ci.org/adamhsparks/nasapower)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/adamhsparks/nasapower?branch=master&svg=true)](https://ci.appveyor.com/project/adamhsparks/nasapower)
[![Coverage Status](https://img.shields.io/codecov/c/github/adamhsparks/nasapower/master.svg)](https://codecov.io/github/adamhsparks/nasapower?branch=master)
[![DOI](https://zenodo.org/badge/109224461.svg)](https://zenodo.org/badge/latestdoi/109224461)

## Introduction

_nasapower_ aims to make it quick, easy to automate downloading
[NASA-POWER](https://power.larc.nasa.gov) agroclimatology data in your R session
as a tidy data frame for agricultural analysis and use in modelling or other
purposes. POWER (Prediction Of Worldwide Energy Resource) data are freely
available for download through a web interface at a resolution of 1˚ longitude
by 1˚ latitude.

Please see
<https://power.larc.nasa.gov/cgi-bin/agro.cgi?email=agroclim@larc.nasa.gov>
for the agroclimatology data web interface.

### Quick start

_nasapower_ is not available from CRAN, only GitHub. It can easily be installed
using the following code:

```{r, eval=FALSE}
# install.packages("devtools", dep = TRUE)
devtools::install_github("adamhsparks/nasapower")
```

_nasapower_ only provides one function, `get_nasa()`, which will download
specified variables for a given 1˚ longitude by 1˚ latitude cell and return a
tidy data frame (long format) of the requested data.

### Usage

To start using `get_nasa()` with the default values, all that is necessary is
to supply the latitude and longitude values. Weather variables will default to 
`T2M`, `T2MN`, `T2MX` and `RH2M` (see below for more weather variables and
their definitions). Start date will default to `1983-1-1` and the end date will
default to the current (system) date.

```{r basic_example, eval=FALSE}
get_nasa(lonlat = c(-89.5, -179.5))
```

Valid `vars` include:

* toa_dwn - Average top-of-atmosphere insolation (MJ/m^2/day)

* swv_dwn - Average insolation incident on a horizontal surface (MJ/m^2/day)

* lwv_dwn - Average downward longwave radiative flux (MJ/m^2/day)

* T2M - Average air temperature at 2m above the surface of the Earth (degrees C)

* T2MN - Minimum air temperature at 2m above the surface of the Earth (degrees C)

* T2MX - Maximum air temperature at 2m above the surface of the Earth (degrees C)

* RH2M - Relative humidity at 2m above the surface of the Earth (%)

* DFP2M - Dew/Frost point temperature at 2m above the surface of the Earth (degrees C)

* RAIN - Average precipitation (mm/day)

* WS10M - Wind speed at 10m above the surface of the Earth (m/s)

## Use of POWER Data

While _nasapower_ does not redistribute the data or provide it in anyway, we
encourage users to follow the requests of the POWER Project Team.

> When POWER data products are used in a publication, we request the following
acknowledgment be included: "These data were obtained from the NASA Langley
Research Center POWER Project funded through the NASA Earth Science Directorate
Applied Science Program."

## Meta

* Please [report any issues or bugs](https://github.com/adamhsparks/nasapower/issues).
* License: MIT
* Get citation information for `nasapower` in R doing `citation(package = "nasapower")`
* Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md).
By participating in this project you agree to abide by its terms.

## References
<https://power.larc.nasa.gov>

<https://power.larc.nasa.gov/documents/Agroclimatology_Methodology.pdf>
