---
title: "Fetch NASA-POWER Parameters"
author: "Adam H Sparks"
date: "`r format(Sys.Date())`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Note on 2019-02-02

_From discussion with the POWER team, this file is apparently for internal use
only and the use for validation as I'm doing here is not its intended use.
However, I do not have a good way to validate requests before sending them to
the API as I am currently awaiting a response from the POWER team on how to
properly validate requests._

In the meantime I have implemented a small fix for the WS2M data in AG while I
await the response from the POWER team.

# Create parameters list for internal checks

These data are used for internal checks to be sure that data requested from the
POWER dataset are valid. The POWER list of parameters that can be queried is
available as a JSON file. Thanks to [raymondben](https://github.com/raymondben)
for pointing me to this file.

## Fetch list from JSON file

Using `jsonlite` read the JSON file into R creating a list.

```{r fetch-JSON}
parameters <-
  jsonlite::fromJSON(
    "https://power.larc.nasa.gov/RADAPP/GEODATA/powerWeb/POWER_Parameters_v110.json"
  )
```

Replace UTF-8 characters in the dataset since R doesn't like this in packages.

```{r, replace-utf}
parameters$SG_DEC_AVG$climatology_definition <-
  gsub("°",
       " degrees",
       parameters$SG_DEC_AVG$climatology_definition)

parameters$SG_HR_SET_ANG$climatology_definition <-
  gsub("°",
       " degrees",
       parameters$SG_HR_SET_ANG$climatology_definition)

parameters$SG_NOON$climatology_definition <-
  gsub("°",
       " degrees",
       parameters$SG_NOON$climatology_definition)
```

## Fix the WS2M issue in AG

See [Issue 32](https://github.com/ropensci/nasapower/issues/32) for more on
this until POWER properly addresses this, we fix it ourselves here.

```{r fix-ws2m}
parameters$WS2M$community <- c("AG", "SB", "SSE")
```

## View list of parameters

The following list has the format:
```
## $PARAMETER_NAME
## [1] "standard_name"
```

Where `PARAMETER_NAME` is used in the internal `parameters` list. The
`"standard_name"` is a longer, more descriptive name for the parameter that may
be more instructive to users.

```{r view-list}
purrr::map(parameters, "standard_name")
```

## Save list for use in `nasapower` package

Using `usethis` save the list as an R data object for use in the
`nasapower` package.

```{r save-list, message=FALSE}
usethis::use_data(parameters, overwrite = TRUE)
```

## Session Info

```{r session-info}
sessioninfo::session_info()
```
