---
title: "Fetch NASA-POWER Parameters"
author: "Adam H Sparks"
date: "`r format(Sys.Date())`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Create `parameters` list for internal checks before sending queries to POWER server

These data are used for internal checks to be sure that data requested from the
POWER dataset are valid. The POWER list of parameters that can be queried is
available as a JSON file. Thanks to [raymondben](https://github.com/raymondben)
for pointing me to this file.

## POWER JSON file

Using `jsonlite` read the JSON file into R creating a list.

```{r fetch-JSON}
parameters <-
  jsonlite::fromJSON(
    "https://power.larc.nasa.gov/RADAPP/GEODATA/powerWeb/POWER_Parameters_v110.json"
  )
```

Replace UTF-8 characters in the dataset since R doesn't like this in packages.

```{r, replace-utf}
parameters$SG_DEC_AVG$climatology_definition <-
  gsub("°",
       " degrees",
       parameters$SG_DEC_AVG$climatology_definition)

parameters$SG_HR_SET_ANG$climatology_definition <-
  gsub("°",
       " degrees",
       parameters$SG_HR_SET_ANG$climatology_definition)

parameters$SG_NOON$climatology_definition <-
  gsub("°",
       " degrees",
       parameters$SG_NOON$climatology_definition)
```

## View list of parameters and units

The following list has the format:
```
## $PARAMETER_NAME
## [1] "standard_name"
```

Where `PARAMETER_NAME` is used in the internal `parameters` list. The
`"standard_name"` is a longer, more descriptive name for the parameter that may
be more instructive to users.

```{r view-list-names}
purrr::map(parameters, "standard_name")
```

The following list has the format:
```
## $PARAMETER_NAME
## [1] AG_Units
## [2] SB_Units
## [3] SSE_Units
```

```{r view-list-units}
purrr::map(parameters, `[`, c("AG_Units", "SB_Units", "SSE_Units"))
```

## Save list for use in `nasapower` package

Using `usethis` save the list as an R data object for use in the
`nasapower` package.

```{r save-list, message=FALSE}
usethis::use_data(parameters, overwrite = TRUE)
```

## Session Info

```{r session-info}
sessioninfo::session_info()
```
