---
title: "Fetch NASA-POWER Parameters"
author: "Adam H Sparks"
date: "`r format(Sys.Date())`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Create parameters list for internal checks

These data are used for internal checks to be sure that data requested from the
POWER dataset are valid. The POWER list of parameters that can be queried is
available as a JSON file. Thanks to [raymondben](https://github.com/raymondben)
for pointing me to this file.

## Fetch list from JSON file

Using `jsonlite` read the JSON file into R creating a list.

```{r fetch_JSON}
parameters <-
  jsonlite::fromJSON(
    "https://power.larc.nasa.gov/RADAPP/GEODATA/powerWeb/POWER_Parameters_v108.json"
  )
```

Replace UTF-8 characters in the dataset since R doesn't like this in packages.

```{r, replace_utf}
parameters$SG_DEC_AVG$climatology_definition <-
  gsub("°",
       " degrees",
       parameters$SG_DEC_AVG$climatology_definition)

parameters$SG_HR_SET_ANG$climatology_definition <-
  gsub("°",
       " degrees",
       parameters$SG_HR_SET_ANG$climatology_definition)


parameters$SG_NOON$climatology_definition <-
  gsub("°",
       " degrees",
       parameters$SG_NOON$climatology_definition)
```

## Save list for use in `nasapower` package

Using `devtools` to save the list as an R data object for use in the
`nasapower` package.

```{r save_list, message=FALSE}
devtools::use_data(parameters, overwrite = TRUE)
```

## Session Info

```{r session_info}
sessioninfo::session_info()
```
