---
title: "nasapower"
author: "Adam H Sparks"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{nasapower}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
LOCAL <- identical(Sys.getenv("LOCAL"), "true")
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>",
                      fig.width = 7,
                      fig.height = 7,
                      fig.align = "center",
		      purl = LOCAL)
```

## Introduction

_nasapower_ aims to make it quick and easy to automate downloading
NASA [POWER](https://power.larc.nasa.gov) global meteorology, surface solar
energy amd climatology data  data in your R session as a tidy data frame for
analysis and use in modelling or other purposes using `get_power()`. POWER
(Prediction Of Worldwide Energy Resource) data are freely available for
download through a web interface at a resolution of 1/2 arc degree longitude
by 1/2 arc degree latitude. Two other functions are provided to quickly and
easily generate weather.met files from the POWER data for use in the
Agricultural Production Systems sIMulator (APSIM), `create_met()` and 
International Consortium for Agricultural Systems Applications (ICASA) format
text files for use in the Decision Support System for Agrotechnology Transfer
(DSSAT) framework, `create_met()`.

## Using `get_power` to Fetch POWER Data

The `get_power()` function has five arguments as seen in this example and
returns a data frame with a metadata header in the current R session.

### Example fetching daily data for a single point

Fetch daily "AG" community temperature, relative humidity and precipitation
for January 1985 for Kingsthorpe, Queensland, Australia.

```{r get-ag, eval=LOCAL}
library(nasapower)
daily_single_ag <- get_power(community = "AG",
                             lonlat = c(151.81, -27.48),
                             pars = c("RH2M", "T2M", "PRECTOT"),
                             dates = c("1985-01-01", "1985-01-31"),
                             temporal_average = "DAILY"
                            )

daily_single_ag
```

### Example fetching daily data for an area

Fetch daily "AG" community relative humidity and temperature for south east
Queensland region.

```{r daily-region, eval=LOCAL}
daily_region_ag <- get_power(community = "AG",
                             lonlat = c(150.5, -28.5 , 153.5, -25.5),
                             pars = c("RH2M", "T2M"),
                             dates = c("1985-01-01", "1985-01-02"),
                             temporal_average = "DAILY")

daily_region_ag
```

### Example fetching interannual data for an area

Fetch interannual solar cooking parameters for south east Queensland region.

```{r interannual, eval=LOCAL}
interannual_sse <- get_power(community = "SSE",
                             lonlat = c(150.5, -28.5 , 153.5, -25.5),
                             dates = c("1984", "1985"),
                             temporal_average = "INTERANNUAL",
                             pars = c("CLRSKY_SFC_SW_DWN",
                                      "ALLSKY_SFC_SW_DWN")
                            )

interannual_sse
```

### Example fetching climatology data

Global data are only available for the climatology `temporal_average`, setting
this argument as such will fetch global data. 

Fetch global "AG"" climatology for temperature and relative
humidity.

```{r get-climatology, eval=LOCAL}
climatology_ag <- get_power(community = "AG",
                            pars = c("T2M", "RH2M"),
                            temporal_average = "CLIMATOLOGY"
                           )

climatology_ag
```

_Note_ the associated metadata in the data frame header are not saved if the
data are exported to a file format other than an R data format, _e.g._, .Rdata,
.rda or .rds.

### Description of parameters for `get_power`

- `community`, a text string with valid values of: "AG" (Agroclimatology), "SSE"
(Surface meteorology and Solar Energy) or "SB" (Sustainable Buildings). The
selected user community will affect the units of the parameter and the temporal
display of time series data (_e.g._, "AG" will use MJ/m^2^/day for radiation
units, while SSE and SB use kW/m^2^/day as units).

- `lonlat`, A numeric vector of geographic coordinates for a cell or region
  entered as x, y coordinates. Not used when `temporal_average` is set to
  "CLIMATOLOGY".

    - *For a single point* To get a specific cell, 1/2 x 1/2 degree, supply a
    length-two numeric vector giving the decimal degree longitude and latitude
    in that order for data to download,  *e.g.*, `lonlat = c(-89.5, -179.5)`.  
  
    - *For regional coverage* To get a region, supply a length-four numeric
    vector as lower left (lon, lat) and upper right (lon, lat) coordinates,
    *e.g.*, `lonlat = c(xmin, ymin, xmax, ymax)` in that order for a given
    region, *e.g.*, a bounding box for the south-western corner of Australia:
    `lonlat = c(112.5, -55.5, 115.5, -50.5)`. *Max bounding box is 10 x 10
    degrees* of 1/2 x 1/2 degree data, *i.e.*, 100 points maximum in total.  
    
  `pars`. A character vector of parameters to download. See `help(parameters)`
  for a full list of valid values and definitions.  If downloading 
  "CLIMATOLOGY" a maximum of 3 `pars` can be specified at one time, for
  "DAILY" and "INTERANNUAL" a maximum of 20 can be specified at one time.

- `dates`, a vector of start and end dates for which to query the POWER
  API, *e.g.*, `dates = c("1983-01-01", "2017-12-31")`. 

- `temporal_average`, a character vector of the desired temporal average(s).
  Valid values are "DAILY", "INTERANNUAL" and "CLIMATOLOGY".

## Using `create_met`

The `create_met()` function wraps `get_power()`, `prepareMet()` and 
`writeMet()` from the [_APSIM_](https://CRAN.R-project.org/package=APSIM)
package to simplify creating .met files from POWER data for use in the
APSIM modelling framework.

Four arguments are passed to this function, `lonlat`, `dates`, `dsn` and
`file_out`.

- `lonlat`

    - *For a single point* To get a specific cell, 1/2 x 1/2 degree, supply a
    length-two numeric vector giving the decimal degree longitude and latitude
    in that order for data to download,  *e.g.*, `lonlat = c(-179.5, -89.5)`.  
  
    - *For regional coverage* To get a region, supply a length-four numeric
    vector as lower left (lon, lat) and upper right (lon, lat) coordinates,
    *e.g.*, `lonlat = c(xmin, ymin, xmax, ymax)` in that order for a given
    region, *e.g.*, a bounding box for the south-western corner of Australia:
    `lonlat = c(112.5, -55.5, 115.5, -50.5)`. *Max bounding box is 10 x 10
    degrees* of 1/2 x 1/2 degree data, *i.e.*, 100 points maximum in total.  
    
- `dates`, a vector of start and end dates for which to query the POWER
  API, *e.g.*, `dates = c("1983-01-01", "2017-12-31")`.

- `dsn`, a file path to the directory for writing the resulting file,
  _.e.g._, "~/Documents". If none is given, defaults to user's home directory.
    
- `file_out`, a file name for the resulting text file written to disk.
   If none is provided, defaults to "APSIM.met"
   
Get POWER values for a single point, Kingsthorpe, Queensland for 1985 and
create an APSIM weather.met object suitable for use in APSIM for crop modelling
and save it to local disk for use in `APSIM` modelling.

```{r create-met, eval=LOCAL}
create_met(lonlat = c(151.81, -27.48),
           dates = c("1985-01-01", "1985-12-31"),
           dsn = tempdir(),
           file_out = "Kingsthorpe.met")
```

## Using `create_icasa`

The `create_icasa()` simplifies a `get_power()` query to create an ICASA file
for use in DSSAT modelling.

Four arguments are passed to this function, `lonlat`, `dates`, `dsn` and
`file_out`.

- `lonlat`

    - *For a single point* To get a specific cell, 1/2 x 1/2 degree, supply a
    length-two numeric vector giving the decimal degree longitude and latitude
    in that order for data to download,  *e.g.*, `lonlat = c(-179.5, -89.5)`.
  
    - *For regional coverage* To get a region, supply a length-four numeric
    vector as lower left (lon, lat) and upper right (lon, lat) coordinates,
    *e.g.*, `lonlat = c(xmin, ymin, xmax, ymax)` in that order for a given
    region, *e.g.*, a bounding box for the south-western corner of Australia:
    `lonlat = c(112.5, -55.5, 115.5, -50.5)`. *Max bounding box is 10 x 10
    degrees* of 1/2 x 1/2 degree data, *i.e.*, 100 points maximum in total.
    
- `dates`, a vector of start and end dates for which to query the POWER
  API, *e.g.*, `dates = c("1983-01-01", "2017-12-31")`.
  
- `dsn`, a file path to the directory for writing the resulting file,
  _.e.g._, "~/Documents". If none is given, defaults to user's home directory.
    
- `file_out`, a  file name for the resulting text file written to disk.
   If none is provided, defaults to "ICASA.txt"

Get POWER values for a single point, Kingsthorpe, Queensland for 1985 and
create an ICASA text file suitable for use in `DSSAT` for crop modelling and
save it to local disk.

```{r create-icasa, eval=LOCAL}
create_icasa(lonlat = c(151.81, -27.48),
             dates = c("1985-01-01", "1985-12-31"),
             dsn = tempdir(),
             file_out = "Kingsthorpe_ICASA.txt")
```

## Creating Spatial Objects from `get_power`

If you require spatial objects to work with, it is rather simple to convert
the resulting tidy data frame from `get_power()` to a spatial object in R using
`raster::rasterFromXYZ()`.

### Converting Regional Data to a _raster_ Object

Using `lapply` and `split` it is easy to create stacks of raster objects from
a regional query.

```{r raster-example, eval=LOCAL}
library(raster)
# Use split to create a list of data frames split by YYYYMMDD
daily_region_ag <- split(daily_region_ag, daily_region_ag$YYYYMMDD)

# Remove date information from data frame, list names will carry YYYYMMDD
daily_region_ag <-
  lapply(daily_region_ag, function(x)
    x[(!names(x) %in% c("YEAR", "MM", "DD", "DOY", "YYYYMMDD"))])

# Create a list of raster bricks from each YYYYMMDD data frame
raster_list <- lapply(daily_region_ag, rasterFromXYZ,
                      crs = "+proj=eqc +lat_ts=0 +lat_0=0 +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")

stack_names <- paste0(names(raster_list), rep(c("_RH2M", "_T2M"), 2))

raster_stack <- stack(unlist(raster_list))
names(raster_stack) <- stack_names
```

Plot all the items in the stack.

```{r plot-fig1, eval=LOCAL}
plot(raster_stack)
```

### Converting Global Climatology to a _raster_ Object

Converting global climatology to a `raster` objects is as simple as querying
and then converting the resulting `tibble` to a _raster_ object using `raster::rasterFromXYZ()`<sup>*</sup>. Illustrated here using the `climatology_ag` object previously created are two methods.

<sub>*: You may experience issues if you have <16GB RAM available converting the entire object to a `raster::brick()`.</sub>

#### For low-memory systems

For low-memory systems (e.g. a recent Apple Mac computer with 8GB RAM), it is useful to have as much of your R global environment free from objects that take much memory and make sure other processes are not using system memory.

Using indexing pick the month or annual value and create a raster object.

```{r create-single-raster, eval=LOCAL}
# split data frame into data frames for each parameter in a list
climatology_list <- split(climatology_ag, climatology_ag$PARAMETER)

# Remove PARAMETER information from data frame, list names supply this
climatology_list <-
  lapply(climatology_list, function(x) {x["PARAMETER"] <- NULL; x})

RH2M_ann <- rasterFromXYZ(climatology_list$RH2M[c(1:2, 15)],
                         crs = "+proj=eqc +lat_ts=0 +lat_0=0 +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")

RH2M_ann

plot(RH2M_ann)
```

#### For systems with more available memory

For systems with RAM of ~16GB or more, it is possible to create a `raster::brick()` of the resulting POWER query. To do so, subset the `tibble()` for the desired parameter and drop the `PARAMETER` column.

```{r create-climatology-brick, eval=FALSE}
T2M_brick <- rasterFromXYZ(subset(climatology_ag, PARAMETER == "T2M")[-3])
```

## References

<https://power.larc.nasa.gov>

<https://power.larc.nasa.gov/documents/POWER_Data_v9_methodology.pdf>

