---
title: "nasapower"
author: "Adam H Sparks"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{nasapower}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>",
                      fig.width = 7,
                      fig.height = 7,
                      fig.align = "center",
                      eval = FALSE)
```

## Introduction

_nasapower_ aims to make it quick and easy to automate downloading
NASA [POWER](https://power.larc.nasa.gov) global meteorology and surface solar
energy climatology data  data in your R session as a tidy data frame for
analysis and use in modelling or other purposes, `get_power()`. POWER
(Prediction Of Worldwide Energy Resource) data are freely available for
download through a web interface at a resolution of 0.5 arc degree longitude
by 0.5 arc degree latitude. Two other functions are provided to quickly and
easily generate weather.met files from the POWER data for use in the
`Agricultural Production Systems sIMulator (APSIM)`, `create_met()` and ICASA
format text files for use in `Decision Support System for Agrotechnology
Transfer (DSSAT)` modelling, `create_met()`.

## Using `get_power`

The `get_power()` function has five arguments as seen in this example and
returns a data frame with a metadata header in the current R session.

### Example fetching daily data for a single point

Fetch daily "AG" community temperature, relative humidity and precipitation
for January 1985 for Kingsthorpe, Queensland, Australia.

```{r get_ag}
library(nasapower)
daily_single_ag <- get_power(community = "AG",
                             lonlat = c(151.81, -27.48),
                             pars = c("RH2M", "T2M", "PRECTOT"),
                             dates = c("1985-01-01", "1985-01-31"),
                             temporal_average = "DAILY"
                            )

daily_single_ag
```

### Example fetching daily data for an area

Fetch daily "AG" community relative humidity and temperature for south east
Queensland region.

```{r daily_region}
daily_region_ag <- get_power(community = "AG",
                             lonlat = c(150.5, -28.5 , 153.5, -25.5),
                             pars = c("RH2M", "T2M"),
                             dates = c("1985-01-01", "1985-01-02"),
                             temporal_average = "DAILY")

daily_region_ag
```

### Example fetching interannual data for an area

Fetch interannual solar cooking parameters for south east Queensland region.

```{r interannual}
interannual_sse <- get_power(community = "SSE",
                             lonlat = c(150.5, -28.5 , 153.5, -25.5),
                             dates = c("1984", "1985"),
                             temporal_average = "INTERANNUAL",
                             pars = c("CLRSKY_SFC_SW_DWN",
                                      "ALLSKY_SFC_SW_DWN")
                            )

interannual_sse
```

### Example fetching climatology data

Fetch global AG climatology for temperature.

```{r get_climatology}
climatology_ag <- get_power(community = "AG",
                            pars = c("T2M", "PRECTOT", "RH2M"),
                            temporal_average = "CLIMATOLOGY"
                           )

climatology_ag
```

_Note_ the associated metadata are not saved if the data are exported to a
file format other than an R data format, _e.g._, .Rdata, .rda or .rds.

### Description of parameters for `get_power`

- `community`, a text string with valid values of: "AG" (Agroclimatology), "SSE"
(Surface meteorology and Solar Energy) or "SB" (Sustainable Buildings). The
selected user community will affect the units of the parameter and the temporal
display of time series data (_e.g._ Agroclimatology will use MJ/m^2^/day for
radiation units, while SSE and SB use kW/m^2^/day as units).

- `lonlat`

    - *For a single point* To get a specific cell, 1/2 x 1/2 degree, supply a
    length-2 numeric vector giving the decimal degree longitude and latitude in
    that order for data to download,  *e.g.*, `lonlat = c(-89.5, -179.5)`.  
  
    - *For regional coverage* To get a region, supply a length-4 numeric
    vector as lower left (lon, lat) and upper right (lon, lat) coordinates,
    *e.g.*, `lonlat = c(xmin, ymin, xmax, ymax)` in that order for a given
    region, *e.g.*, a bounding box for the south-western corner of Australia:
    `lonlat = c(112.5, -55.5, 115.5, -50.5)`. *Max bounding box is 10 x 10
    degrees* of 1/2 x 1/2 degree data, *i.e.*, 100 points maximum in total.  
  
    - *For global coverage* To get global coverage for long term monthly
    averages for the entire globe use `Global` in place of `lonlat` values.
    `temporal_average` will automatically be set to `CLIMATOLOGY` if this option
    is set.
    
- `pars`, A character vector of solar, meteorological or climatology
  parameters to download.  See `names(parameters)` for a full list of
  valid values and definitions.  Visit the 
  [POWER website](https://power.larc.nasa.gov/#paramstable) for the Parameters
  Table. If downloading CLIMATOLOGY a maximum of 3 `pars` can be
  specified at one time for for DAILY and INTERANNUAL a maximum of 20 can
  be specified at one time.

- `dates`, a vector of start and end dates for which to query the POWER
  API, *e.g.*, `dates = c("1983-01-01", "2017-12-31")`. 

- `temporal_average`, a character vector of the desired temporal average(s).
  Valid values are "DAILY", "INTERANNUAL" and "CLIMATOLOGY".

## Using `create_met`

The `create_met()` function wraps `get_power()`, `prepareMet()` and 
`writeMet()` from the [_APSIM_](https://CRAN.R-project.org/package=APSIM)
package to simplify creating .met files from POWER data for use in the
APSIM modelling framework.

Four arguments are passed to this function, `lonlat`, `dates`, `dsn` and
`file_out`.

- `lonlat`

    - *For a single point* To get a specific cell, 1/2 x 1/2 degree, supply a
    length-2 numeric vector giving the decimal degree longitude and latitude in
    that order for data to download,  *e.g.*, `lonlat = c(-179.5, -89.5)`.  
  
    - *For regional coverage* To get a region, supply a length-4 numeric
    vector as lower left (lon, lat) and upper right (lon, lat) coordinates,
    *e.g.*, `lonlat = c(xmin, ymin, xmax, ymax)` in that order for a given
    region, *e.g.*, a bounding box for the south-western corner of Australia:
    `lonlat = c(112.5, -55.5, 115.5, -50.5)`. *Max bounding box is 10 x 10
    degrees* of 1/2 x 1/2 degree data, *i.e.*, 100 points maximum in total.  
    
- `dates`, a vector of start and end dates for which to query the POWER
  API, *e.g.*, `dates = c("1983-01-01", "2017-12-31")`.

- `dsn`, a file path to the directory for writing the resulting file,
  _.e.g._, "~/Documents". If none is given, defaults to user's home directory.
    
- `file_out`, a file name for the resulting text file written to disk.
   If none is provided, defaults to "APSIM.met"
   
Get POWER values for a single point, Kingsthorpe, Queensland for 1985 and
create an APSIM weather.met object suitable for use in APSIM for crop modelling
and save it to local disk for use in `APSIM` modelling.

```{r create_met}
create_met(lonlat = c(151.81, -27.48),
           dates = c("1985-01-01", "1985-12-31"),
           dsn = tempdir(),
           file_out = "Kingsthorpe.met")
```

## Using `create_icasa`

The `create_icasa()` simplifies a `get_power()` query to create an ICASA file
for use in `DSSAT` modelling.

Four arguments are passed to this function, `lonlat`, `dates`, `dsn` and
`file_out`.

- `lonlat`

    - *For a single point* To get a specific cell, 1/2 x 1/2 degree, supply a
    length-2 numeric vector giving the decimal degree longitude and latitude in
    that order for data to download,  *e.g.*, `lonlat = c(-179.5, -89.5)`.
  
    - *For regional coverage* To get a region, supply a length-4 numeric
    vector as lower left (lon, lat) and upper right (lon, lat) coordinates,
    *e.g.*, `lonlat = c(xmin, ymin, xmax, ymax)` in that order for a given
    region, *e.g.*, a bounding box for the south-western corner of Australia:
    `lonlat = c(112.5, -55.5, 115.5, -50.5)`. *Max bounding box is 10 x 10
    degrees* of 1/2 x 1/2 degree data, *i.e.*, 100 points maximum in total.
    
- `dates`, a vector of start and end dates for which to query the POWER
  API, *e.g.*, `dates = c("1983-01-01", "2017-12-31")`.
  
- `dsn`, a file path to the directory for writing the resulting file,
  _.e.g._, "~/Documents". If none is given, defaults to user's home directory.
    
- `file_out`, a  file name for the resulting text file written to disk.
   If none is provided, defaults to "ICASA.txt"

Get POWER values for a single point, Kingsthorpe, Queensland for 1985 and
create an ICASA text file suitable for use in `DSSAT` for crop modelling and
save it to local disk.

```{r create_icasa}
create_icasa(lonlat = c(151.81, -27.48),
             dates = c("1985-01-01", "1985-12-31"),
             dsn = tempdir(),
             file_out = "Kingsthorpe_ICASA.txt")
```

## Creating Spatial Objects from `get_power`

If you require spatial objects to work with, it is rather simple to convert
the resulting tidy data frame from `get_power()` to a spatial object in R using `raster::rasterFromXYZ()`.

### Converting Regional Data to a _raster_ Object

Using `lapply` and `split` it is easy to create stacks of raster objects from
a regional query.

```{r raster_example}
library(raster)
# Use split to create a list of data frames split by YYYYMMDD
daily_region_ag <- split(daily_region_ag, daily_region_ag$YYYYMMDD)

# Remove date information from data frame, list names will carry YYYYMMDD
daily_region_ag <-
  lapply(daily_region_ag, function(x)
    x[(!names(x) %in% c("YEAR", "MM", "DD", "DOY", "YYYYMMDD"))])

# Create a list of raster bricks from each YYYYMMDD data frame
raster_list <- lapply(daily_region_ag, rasterFromXYZ,
                      crs = "+proj=eqc +lat_ts=0 +lat_0=0 +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ")

stack_names <- paste0(names(raster_list), rep(c("_RH2M", "_T2M"), 2))

raster_stack <- stack(unlist(raster_list))
names(raster_stack) <- stack_names
```

Plot all the items in the stack.

```{r plot_fig1}
plot(raster_stack)
```

### Converting Global Climatology to a _raster_ Object

Converting global climatology to a `raster::brick` is as simple as querying
and then converting the resulting `tibble` to a raster object. Illustrated here
using the `climatology_ag` object previously created.

```{r global_climatology}
# create RasterBricks for the individual parameters and drop the parameter field
RH2M <- rasterFromXYZ(subset(climatology_ag, PARAMETER == "RH2M")[, -3])

T2M <- rasterFromXYZ(subset(climatology_ag, PARAMETER == "T2M")[, -3])

PRECIP <- rasterFromXYZ(subset(climatology_ag, PARAMETER == "PRECTOT")[, -3])
```

Plot the annual values from the resulting `RasterBrick()` objects.

```{r plot_climatology}
plot(RH2M$ANN)

plot(T2M$ANN)

plot(PRECIP$ANN)
```


## References

<https://power.larc.nasa.gov>

<https://power.larc.nasa.gov/documents/POWER_Data_v8_methodology.pdf>
