---
title: "nasapower"
author: "Adam H Sparks"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{nasapower}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(nasapower)
knitr::opts_chunk$set(eval = FALSE,
                      collapse = TRUE,
                      comment = "#>",
                      fig.width = 7,
                      fig.height = 7,
                      fig.align = "center")
```

## Introduction

_nasapower_ aims to make it quick, easy and efficient to automate downloading
NASA-POWER agroclimatology data in your R session as a tidy data frame.

_nasapower_ only provides one function, `get_power()`, which will download
specified variables and return a tidy data frame of the requested data. Weather
variables can be specified by using the `pars` argument. 

## Using get_power()

The `get_power()` function requires five arguments as seen in this example.

```{r example, eval=FALSE}
power <- get_power(community = "AG",
                   latlon = c(-179.5, -89.5),
                   pars = c("RH2M", "T2M"),
                   dates = c("1985-01-01", "1985-12-31"),
                   temporal_average = "daily")
```

- `community`, a text string with valid values of: "AG" (Agroclimatology), "SSE"
(Surface meteorology and Solar Energy) or "SB" (Sustainable Buildings). The
selected user community will affect the units of the parameter and the temporal
display of time series data (_e.g._ Agroclimatology will use MJ/m^2^/day for
radiation units, while SSE and SB use kW/m^2^/day as units).

- `latlon`, a length-2 numeric vector giving the decimal degree longitude and
and latitude coordinates in that order for cell data to download or a length-4
numeric vector giving the decimal degree longitude and and latitude coordinates
forming a bounding box as ymin, xmin, ymax, xmax in that order. When specifying
a region, the maximum amount of cells that can be downloaded are 100, i.e., a 10
x 10 0.5˚ bounding box or 5 x 5 1˚ bounding box. Leave blank for global
coverage.

- `pars`, A character vector of solar or meteorological variables to download.
See the `Value` field of [`parameters`](../help/parameters) for a full list of
valid values or visit the [POWER website](https://power.larc.nasa.gov/new/) for
the Parameter Dictionary. If downloading global coverage a maximum of 3 `pars`
can be specified at one time.

- `dates`, a vector of start and end dates for which to query the NASA-POWER
API, *e.g.*, `dates = c("1983-01-01", "2017-12-31")`. 

- `temporal_average`, a character vector of the desired temporal average(s).
Valid values are "DAILY", "INTERANNUAL" and "CLIMATOLOGY".

## Creating Spatial Objects

If you require spatial objects to work with, it is rather simple to convert
the resulting tidy data frame from `get_power()` to a spatial object in R using `raster::rasterFromXYZ()`.

### Converting to a _raster_ Object

Using the `example` object from the previous example, convert the data frame to a
raster brick. The LON and LAT values are in columns 6 and 7 with the weather
vars in any following columns.

The first step is to `base::split()` the data frame into a list of data frames by day.

Then convert the `base::list()` into a `list` of `raster::brick()` objects.

```{r raster_example}
library(raster)

# Use split to create a list of data frames split by YYYYMMDD
PHL <- split(PHL, PHL$YYYYMMDD)

# Remove date information from data frame, list names will carry YYYYMMDD
PHL <-
  lapply(PHL, function(x)
  x[(!names(x) %in% c("YEAR", "MONTH", "DAY", "YYYYMMDD", "DOY"))])

# Create a list of raster bricks from each YYYYMMDD data frame
PHL <- lapply(PHL, rasterFromXYZ,
              crs = "+proj=eqc +lat_ts=0 +lat_0=0 +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ")
```

## References

<https://power.larc.nasa.gov>

<https://power.larc.nasa.gov/documents/Agroclimatology_Methodology.pdf>
