---
title: "Use Case: Using nasapower and the APSIM R Package to Create APSIM .met Files"
author: "Adam H. Sparks
"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r install_apsim, eval=FALSE}
install.packages("APSIM")
```

Load the packages necessary.

```{r load_libraries}
library("APSIM")
library("nasapower")
library("rnaturalearth")
```

## Get NASA-POWER data

Using the `APSIM::prepareMet()` example where data for Kingsthorpe, Qld,
is provided by the package, we can compare it with NASA-POWER data for
Kingsthorpe, Queensland for 2010.

To do that first we need to use `get_cell()` to fetch the data for 2010 and
create a data frame of weather data which includes the variables, T2M, T2MN,
T2MX, RAIN and WSM10M.

```{r get_data, eval=TRUE}
kings_nasa <- get_cell(
  lonlat = c(151.81, -27.48),
  stdate = "2010-1-1",
  endate = "2010-12-31",
  vars = c("T2M", "T2MN", "T2MX", "RAIN", "WS10M")
  )

kings_nasa
```

Now that we have the data...

## Create the .met file

```{r create_met, eval=TRUE}
# reorder the data columns from nasapower
kings_met <- kings_nasa[...]

# from the APSIM documentation example
new_names <-
  c("Date",
    "maxt",
    "mint",
    "rain",
    "evaporation",
    "radn",
    "vp",
    "Wind",
    "RH",
    "SVP")

units <-
  c("()",
    "(oC)",
    "(oC)",
    "(mm)",
    "(mm)",
    "(MJ/m^2/day)",
    "()",
    "()",
    "()",
    "()")

prepareMet(kings_met,
           lat = kings_nasa[1, "LAT"],
           lon = kings_nasa[1, "LON"],
           newNames = new_names,
           units = units)
```

## Using get_region() with APSIM

To use `get_region()` with `APSIM::prepareMet()` we need to make separate `met`
files for each cell from the resulting data frame object, which is easy to do
using _dplyr_.

```{r get_regions, eval=FALSE}
oz <- get_region(lonlat = c(112.91972, 159.256088, -55.11694, -9.221099),
                    stdate = "2017-1-1", endate = "2017-1-2")
```

```{r create_dfs, eval=FALSE}
library(dplyr)
library(purrr)

# add unique id values (lon-lat value) and sort by that id
oz <- oz %>%
  mutate(ids = paste(LON, LAT, sep = '-')) %>% 
  arrange(id)

# create list of data frames with unique id values
oz_list <- split(oz, oz$id)

# apply the prepareMet() function to the list of data frames
met_list <- map(.x = df2, .f = prepareMet, units = units)
```
