---
title: "Use Case: Using nasapower and the APSIM R Package to Create APSIM .met Files"
author: "Adam H. Sparks
"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
        toc: true
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```

```{r install_apsim, eval=FALSE}
install.packages("APSIM")
```

Load the packages necessary.

```{r load_libraries, message=FALSE}
library(APSIM)
library(nasapower)
library(rnaturalearth)
library(lubridate)
library(ggplot2)
library(dplyr)
```

NASA - POWER data are often used with APSIM (White _et al._ 2012) or other
models where data may be missing or as a whole data set where data are not
readily available, _e.g._ modelling rice disease globally (Savary _et al._
2012).

In this use case we illustrate how to compare NASA-POWER data with station
data and format for use in APSIM using the R package, APSIM (Fainges 2017).

## Get NASA-POWER Data

Using the `APSIM::prepareMet()` example where data for Kingsthorpe, Qld,
is provided by the package, we can compare it with NASA-POWER data for
Kingsthorpe, Queensland for 2010.

To do that first we need to use `get_cell()` to fetch the data for 2010 and
create a data frame of weather data which includes the variables, T2M, T2MN,
T2MX, RAIN and WSM10M.

```{r get_data_show, eval=FALSE}
kings_nasa <- get_cell(
  lonlat = c(151.81, -27.48),
  stdate = "2010-1-1",
  endate = "2010-12-31",
  vars = c("T2MX", "T2MN", "RAIN", "WS10M", "RH2M")
  )
```

```{r get_data_execute, eval=TRUE, message=FALSE, include=FALSE}
# This code chunk actually downloads the data, but it will show
# download progress, so we don't include it.
kings_nasa <- get_cell(
  lonlat = c(151.81, -27.48),
  stdate = "2010-1-1",
  endate = "2010-12-31",
  vars = c("T2MX", "T2MN", "RAIN", "WS10M", "RH2M")
  )
```

## Compare the data

To compare the data more easily using graphs we'll join the two data frames
using _dplyr_. 

First convert the Date column from kingsData to a Date object using _lubridate_.

```{r conv_date, eval=TRUE}
kingsData$Date <- lubridate::dmy(kingsData$Date)
```

Next, join the two data frames together to create a data frame called,
`compare`.

```{r compare, eval=TRUE}
compare <- left_join(kingsData, kings_nasa, by = c("Date" = "YYYYMMDD"))
```

### Graphing the data

Graph maximum temperature, comparing `kingsData` to NASA-POWER data and check
correlation between the values.

### Maximum Temperature

```{r graph_maxt, eval=TRUE}
ggplot(compare, aes(x = max.temp, y = T2MX)) +
  geom_point(colour = "red") +
  theme_minimal()

cor(compare$max.temp, compare$T2MX)
```

### Miniumum Temperature

Graph minimum temperature, comparing `kingsData` to NASA-POWER data.

```{r graph_mint, eval=TRUE}
ggplot(compare, aes(x = min.temp, y = T2MX)) +
  geom_point(colour = "blue") +
  theme_minimal()
  
  cor(compare$min.temp, compare$T2MN)
```

### Rainfall

```{r graph_rain, eval=TRUE}
ggplot(compare, aes(x = rain, y = RAIN)) +
  geom_point(colour = "dark blue") +
  theme_minimal()
  
  cor(compare$rain, compare$RAIN)
```

### Radiation

```{r graph_rad, eval=FALSE}
ggplot(compare, aes(x = rad, y = ???)) +
  geom_point(colour = "orange") +
  theme_minimal()
  
  cor(compare$rad, compare$???)
```


### Relative Humidity

Note that due to the presence of `NA` values in the `RH.at.9am`, the `cor`
function will not work with default arguments, use `pairwise.complete.obs`.

```{r graph_RH, eval=TRUE}
ggplot(compare, aes(x = RH.at.9am, y = RH2M)) +
  geom_point(colour = "dark green") +
  theme_minimal()
  
cor(compare$RH.at.9am, compare$RH2M, use = "pairwise.complete.obs")
```

Now that we have the data...

## Create the .met file

```{r create_met, eval=FALSE}
# reorder the data columns from nasapower
kings_met <- kings_nasa[...]

# from the APSIM documentation example
new_names <-
  c("Date",
    "maxt",
    "mint",
    "rain",
    "evaporation",
    "radn",
    "vp",
    "Wind",
    "RH",
    "SVP")

units <-
  c("()",
    "(oC)",
    "(oC)",
    "(mm)",
    "(mm)",
    "(MJ/m^2/day)",
    "()",
    "()",
    "()",
    "()")

prepareMet(kings_met,
           lat = kings_nasa[1, "LAT"],
           lon = kings_nasa[1, "LON"],
           newNames = new_names,
           units = units)
```

## Using get_region() with APSIM

To use `get_region()` with `APSIM::prepareMet()` we need to make separate `met`
files for each cell from the resulting data frame object, which is easy to do
using _dplyr_.

```{r get_regions, eval=FALSE}
oz <- get_region(lonlat = c(112.91972, 159.256088, -55.11694, -9.221099),
                    stdate = "2017-1-1", endate = "2017-12-31")
```

```{r create_dfs, eval=FALSE}
library(dplyr)

# mutate: add unique id values (lon-lat value)
# arrange: sort data frames by that id
# split: take the data.frame() and split it into a list of data.frames by id
oz_list <- oz %>%
  mutate(id = paste(LON, LAT, sep = '-')) %>%
  arrange(id) %>%
  split(., .$id)

# from the APSIM documentation example
new_names <-
  c("Date",
    "maxt",
    "mint",
    "rain",
    "evaporation",
    "radn",
    "vp",
    "Wind",
    "RH",
    "SVP")

units <-
  c("()",
    "(oC)",
    "(oC)",
    "(mm)",
    "(mm)",
    "(MJ/m^2/day)",
    "()",
    "()",
    "()",
    "()")

# map:: apply the prepareMet() function to the list of data frames
# write each item for the list to a file on disk for use in APSIM
met_list <- lapply(oz_list, function(i)
  prepareMet(oz_list[i],
           lat = oz_list[[i]][1, "LAT"],
           lon = oz_list[[i]][1, "LON"],
           newNames = new_names,
           units = units))

# write met files to disk
lapply(1:length(met_list),
       function(i) writeMetFile(met_list[[i]],
                                fileName = paste0(names(met_list[i]))))
```

## References

Justin Fainges (2017) "APSIM: General Utility Functions for the 'Agricultural
Production Systems Simulator'". R package version 0.9.2.
<https://CRAN.R-project.org/package=APSIM>

Serge Savary, Andrew Nelson, Laetitia Willocquet, Ireneo Pangga and Jorrel
Aunario (2012) "Modeling and mapping potential epidemics of rice diseases
globally" Crop Protection, Volume 34, Pages 6-17, ISSN 0261-2194 DOI:
10.1016/j.cropro.2011.11.009

Jeffrey W. White, Gerrit Hoogenboom, Paul W. Stackhouse, James M. Hoell (2009)
"Evaluation of NASA satellite- and assimilation model-derived long-term daily
temperature data over the continental US" Agricultural and Forest
Meteorology, Volume 148, Issue 10, 2008, Pages 1574-1584, ISSN 0168-1923, DOI:
10.1016/j.agrformet.2008.05.017.
